Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-28 WSSSPE1_report/WSSSPE1_summary_report.tex
more edits

%Summary of above discussed cases: lessons, trends, commonalities, is there an ideal sustainability scenario?
In order to investigate the much discussed issue of career paths for developers
of scientific software, we conducted a small survey of the use-cases discussed
above with respect to the current status of its core developers. We looked at
the workshop submissions and the original cited papers for the respective
software. We looked at the webpages and source codes of the earliest versions
where available to find the original authors. we find that in a majority of
scientific software more than one original developers are still around, in some
cases for more than 10 years, (eg. \emph{R/qtl}). Whereas, in the case of
utility software projects, it appears that at least one original developer is
still around often in a part-time role. In other cases, it appears that the
original developers have went on to assume a larger role, again with the same
project.

From this survey, it appears that the career paths for utility software
developers are broader compared to the developers of niche software. However,
as projects as a whole, the small-scale, specific software have a lower risk
and cost in terms of sustaining as compared to utility software. Often the
sustainability of such projects are achieved by the fact that the core
developer or team heavily utilizes the software for their science, eg.
\emph{R/qtl, PETSc}. The development of scientific software requires more
scientific skills compared with those of utility software. This diminishes the
need of specialist programmers and practices in many cases where the bulk of
development is carried out by a domain scientist.

>>>>

%Summary of above discussed cases: lessons, trends, commonalities, is there an ideal sustainability scenario?
In order to investigate the much discussed issue of career paths for developers
of scientific software, we conducted a small survey of the use-cases discussed
above with respect to the current status of its core developers. We looked at
the workshop submissions and the original cited papers for the respective
software. Where available, we looked at the webpages and source code of the
earliest versions to find the original authors. We excluded relatively new
projects such as \emph{SciPy} and \emph{numFOCUS} from this survey. we find
that in a majority of scientific software more than one original developers are
still around, in some cases for more than 10 years, (eg. \emph{R/qtl}).
Whereas, in the case of utility software projects, it appears that at least one
original developer is still around often in a part-time role. In other cases,
it appears that the original developers have went on to assume a larger role,
again with the same project.

From this survey, it appears that the career paths for utility software
developers are broader compared to the developers of niche software. However,
as projects as a whole, the small-scale, specific software have a lower risk
and cost in terms of sustaining as compared to utility software. Often the
sustainability of such projects is achieved by the fact that the core
developer or team heavily utilizes the software for their science, eg.
\emph{R/qtl, PETSc}. Furthermore, the development of scientific software requires more
scientific skills compared with those of utility software. This diminishes the
need of specialist programmers and practices in many cases where the bulk of
development is carried out by a domain scientist.
